---
title: "module3"
author: "Vishal"
date: "2/17/2022"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document describes dealing with R data structures and using indexing to access and modify the data.

Loads the package 
```{r}
library(svm346)
```

Sets seed to 1
```{r}
set.seed(1)
```

Create random sample of size 20 between 0-100 
```{r}
v1 <- sample(0:100, size = 20, replace = TRUE)
v1
```
Create random sample of size 20 between 0-10 

```{r}
v2 <- sample(1:10, size = 20, replace = TRUE)
v2
```
Create sample of size 20 uniformly distributed between 0-1

```{r}
v3 <- runif(n = 20, min = 0, max = 1)
v3
```
Create vector of letters using values from v2 as index

```{r}
v4 <- vector()
for(i in 1:length(v2)) {
  v4[i] = LETTERS[v2[i]]
}
v4
```

Create matrix m1 using vectors v1 and v2

```{r}
m1 <- cbind(v1, v2)
m1
```
Create matrix m2 using vectors v2 and v3

```{r}
m2 <- cbind(v2, v3)
m2
```

Create matrix m3 using vectore v3 and v4
```{r}
m3 <- cbind(v3, v4)
m3
```

Create dataframe DF using vectors v1, v2, v3, v4
```{r}
DF <- data.frame(v1, v2, v3, v4)
DF
```

Question: why shouldn’t you use the name “df” to name this object?
Answer: Using df might cause unwanted errors because it is predefined in the tree package

Changing column names of the dataframe DF
```{r}
names(DF) <- c("a","b", "c", "grp")
```

Question: why is it okay to rename a column in DF “c”, but not to rename v3 that name?
Answer: v3 is the vector which stores the values. DF is created using v3. changing the column name will just change the reference in the data frame.

```{r}
str(DF)
```

What kind of data type is “grp”? How would you change it between character and factor and back again?

"grp" has a datatype of character. Following code helps convert columns with character to factor data type. and then the factor back to character.

```{r}
DF <- as.data.frame(unclass(DF), stringsAsFactors = TRUE)
str(DF)

DF$grp <- as.character(DF$grp)
str(DF)
DF
```

Adding vectors v1 and v2 using my_calc function
```{r}
s1 <-  my_calc(v1, v2, "sum")
s1
```
Adding vectors v2 and v3 using my_calc function

```{r}
s2 <-  my_calc(v2, v3, "sum")
s2
```

Adding vectors v1 and v3 using my_calc function

```{r}
s3 <-  my_calc(v1, v3, "sum")
s3
```

Adding matrices m1 and m1 using my_calc function

```{r}
s4 <-  my_calc(m1, m1, "sum")
s4
```
Adding matrices m1 and m2 using my_calc function


```{r}
s5 <-  my_calc(m1, m2, "sum")
s5
```
Adding matrices m2 and m3 using my_calc function

This will throw an error because m3 has 2nd column as character data and m2 has 2nd column as numeric data. Binary operations cannot be performed with character data

```{r, error= TRUE}
s6 <-  my_calc(m2, m3, "sum")
```

Dividing vectors v1 and v2 using my_calc function

```{r}
d1 <-  my_calc(v1, v2, "divide")
d1
```

Dividing vectors v2 and v3 using my_calc function

```{r}
d2 <-  my_calc(v2, v3, "divide")
d2
```

Dividing vectors v1 and v3 using my_calc function

```{r}
d3 <-  my_calc(v1, v3, "divide")
d3
```

Dividing matrices m1 and m1 using my_calc function

```{r}
d4 <-  my_calc(m1, m1, "divide")
d4
```

Dividing vectors m1 and m2 using my_calc function

```{r}
d5 <-  my_calc(m1, m2, "divide")
d5
```

Dividing vectors m2 and m3 using my_calc function
This will throw an error because m3 has 2nd column as character data and m2 has 2nd column as numeric data. Binary operations cannot be performed with character data


```{r, error=TRUE}
d6 <-  my_calc(m2, m3, "divide")
```
Multiplying row1, column1 by row1, column2 using my_calc function

```{r}

df_m1 <- my_calc(DF[[1]][[1]], DF[[2]][[1]], "multiply")
df_m1
```
Multiplying rows 1-10, column 1-3 by rows1-10, columns 1-3

```{r}
df1 <- DF[1:10,1:3]
df1
df_m2 <- my_calc(df1, df1, "multiply")
df_m2

```
Filtering rows from DF whose grp column value is "E"

```{r}
filtered_df <- DF[DF$grp == "E", c("a", "b", "c")]
filtered_df

```

Dividing columns a,b by column c of filtered_df using my_calc function

```{r}
df3 <- my_calc(filtered_df[c("a")], filtered_df[c("c")], fun = "divide")
df3


df4 <- my_calc(filtered_df[c("b")], filtered_df[c("c")], fun = "divide")
df4
```

Dividing columns a,b by column c of filtered_df 

```{r}
df4 <- filtered_df$a/filtered_df$c
df4

df5 <- filtered_df$b/filtered_df$c
df5
```

Create list l1 using v1, m1 and m2
```{r}
l1 <- list(v1, m1, m2)
l1

```

Iterates list1 and applies sum function on each item in the list

```{r}
l2 <- lapply(l1, sum)
l2
```

Iterates over list1 and multiplies each item with v2

```{r}
l3 <- lapply(l1, function(x) {
  return(my_calc(x, v2, "multiply"))
})
l3

```

Iterates over l1 and multiplies each item with v2

```{r}
s1 <- sapply(list1, function(x) {
  return(my_calc(x, v2, "multiply"))
})
s1

str(s1)
```











